Whole-body connectome of a segmented annelid larva

### 2024-10-25 Manuscript.qmd
Ms: edits, upload odt w.o. figs

```{r, out.width = "100%", fig.cap="**Figure 2. Modularity of the *Platynereis* larval connectome.** Full connectome graph of the *P. dumerilii* three-day-old larva chemical synapse connectome. Nodes represent individual cells, edges represent synaptic connectivity. Nodes are coloured by modules. Node sizes are proportional to weighted degree. The individual panels show the morphology of all cells in each module with the outline of the yolk and aciculae shown in grey. <br> Figure 2—source data 1. Source data of the connectome graph in tibble graph (tbl_graph) format saved as an R binary object. "}
#| echo: false 
knitr::include_graphics("Figures/Figure2.png")
```

@@

Supplementary Table 1 lists all neuronal and non-neuronal cell types, their main annotations, including morphological, cell-class, segment, ganglion or sensory organ, neurotransmitter and neuropeptide phenotype, and main synaptic partners.
We also listed all literature references for previously published neuronal and non-neuronal cells and indicated which cells are reported here for the first time (105 cell types out of 294).
Interactive 3D morphological renderings of each cell type together with their main annotations can also be explored on a webpage (https://jekelylab.github.io/Platynereis_connectome/celltype_compendium/index.html and Supplementary file 1).
Querying by two or more annotations allows tallying the number of cells in different category (e.g. number of sensory cells with a single penetrating cilium or the number of gland cells in a particular segment)(Figure 3---figure supplement 3).

@@

```{r, out.width = "100%", fig.cap="**Figure 4. Cell-type-level connectivity of the *Platynereis* larval connectome.** (A) The cell-type connectome of the three-day-old larva. Nodes represent grouped cells of the same type, edges represent synaptic connectivity (square root of the sum of synapses). Nodes are coloured by cell class (SN - orange; IN - magenta; MN - blue; effector - grey). (B) Table of cell type and network statistics. (C) Histogram of edge weights. (D) Size of the largest network after removing edges of increasing weights. (E) Number of sensory neurons with different path distances from effector cells. (F) Number of SN, IN, MN and effector cell types in the cell-type-level connectome shown for all nodes, source nodes, sink nodes and cut nodes. (G) Histogram of weighted degree of nodes, plotted for each cell class. (H) Histogram of pagerank of nodes, plotted for each cell class. (I) Weighted degree of nodes in the cell-type connectome in relation to node pagerank-centrality. <br> Figure 4—source data 1. Source data of the cell type connectivity graph in tbl_graph format saved as an R binary file. "}
#| echo: false
knitr::include_graphics("Figures/Figure4.png")
```

@@

```{r, out.width = "70%", fig.cap="**Figure 5. Transmitter phenotypes mapped at single-cell-resolution. ** (A) The cell-type connectome graph with nodes coloured based on neurotransmitter phenotype. (B, C) Morphological rendering of neurons with immunogold labelling for pigment dispersing factor (PDF) neuropeptide, ventral (B) and anterior (C) views. (D) Neurons with immunogold labelling for leucokinin. (E) Neurons with immunogold labelling for allatotropin/orexin. (F) Neurons with immunogold labelling for FVamide. (G) Neurons with immunohistochemically mapped expression of sNPF/RYamide, RGWamide and pedal peptide 2/MLD neuropeptides. (H) Neurons with immunohistochemically mapped expression of achatin and immunogold- or immunohistochemically-mapped (SN-IRP2-FMRF) FMRFamide neuropeptide. (I) Neurons with immunogold (SNMIP-vc) or immunohistochemically mapped expression of myoinhibitory peptide (MIP) expression. (J) Neurons with immunohistochemically mapped expression of serotonin and genetically mapped expression of tryptophan hydroxylase (TrpH), a serotonergic marker. (K) Neurons with genetically mapped expression of vesicular glutamate transporter (VGluT), a glutamatergic marker. (L, M) Neurons with genetically mapped expression of vesicular acetylcholine transporter (VAChT) and choline acetyltransferase (ChAT), cholinergic markers, ventral (L) and anterior (M) views. <br> Figure 5—source data 1. Interactive html file with the network shown in panel A."}
#| echo: false
knitr::include_graphics("Figures/Figure5.png")
```

@@

```{r, out.width = "100%", fig.cap="**Figure 4—figure supplement 6. Grouped connectivity graph. ** Cell-type level connectome graph with node horizontal positions proportional to the relative ratio of incoming (postsynaptic) and outgoing (presynpatic) sites. Cell-type names are shown under the nodes. <br> Figure 4—figure supplement 6—source data 1. The network in visNetwork format saved as an R RDS source file, saved with bz2 compression. Can be loaded with read_rds(). "}
#| echo: false
knitr::include_graphics("Figures/Figure4_fig_suppl6.png")
```

@@

Supplementary file 1. 
A compressed (zip) stand-alone webpage listing all cell types together with their interactive 3D morphological rendering. After uncompressing the file, the index.html file can be opened in a browser.

>>>>

```{r, out.width = "100%", fig.cap="**Figure 2. Modularity of the *Platynereis* larval connectome.** Full connectome graph of the *P. dumerilii* three-day-old larva chemical synapse connectome. Nodes represent individual cells, edges represent synaptic connectivity. Nodes are coloured by modules. Node sizes are proportional to weighted degree. The individual panels show the morphology of all cells in each module with the outline of the yolk and aciculae shown in grey. <br> Figure 2—source data 1. Source data of the connectome graph in tibble graph (tbl_graph) format saved as an R binary object. For an interactive version of the graph see also https://jekelylab.github.io/Platynereis_connectome/."}
#| echo: false 
knitr::include_graphics("Figures/Figure2.png")
```

@@

Supplementary Table 1 lists all neuronal and non-neuronal cell types, their main annotations, including morphological, cell-class, segment, ganglion or sensory organ, neurotransmitter and neuropeptide phenotype, and main synaptic partners.
We also listed all literature references for previously published neuronal and non-neuronal cells and indicated which cells are reported here for the first time (105 cell types out of 294).
Interactive 3D morphological renderings of each cell type together with their main annotations can also be explored on a webpage (https://jekelylab.github.io/Platynereis_connectome/).
Querying by two or more annotations allows tallying the number of cells in different category (e.g. number of sensory cells with a single penetrating cilium or the number of gland cells in a particular segment)(Figure 3---figure supplement 3).

@@

```{r, out.width = "100%", fig.cap="**Figure 4. Cell-type-level connectivity of the *Platynereis* larval connectome.** (A) The cell-type connectome of the three-day-old larva. Nodes represent grouped cells of the same type, edges represent synaptic connectivity (square root of the sum of synapses). Nodes are coloured by cell class (SN - orange; IN - magenta; MN - blue; effector - grey). (B) Table of cell type and network statistics. (C) Histogram of edge weights. (D) Size of the largest network after removing edges of increasing weights. (E) Number of sensory neurons with different path distances from effector cells. (F) Number of SN, IN, MN and effector cell types in the cell-type-level connectome shown for all nodes, source nodes, sink nodes and cut nodes. (G) Histogram of weighted degree of nodes, plotted for each cell class. (H) Histogram of pagerank of nodes, plotted for each cell class. (I) Weighted degree of nodes in the cell-type connectome in relation to node pagerank-centrality. <br> Figure 4—source data 1. Source data of the cell type connectivity graph in tbl_graph format saved as an R binary file. For an interactive version of the graph see also https://jekelylab.github.io/Platynereis_connectome/."}
#| echo: false
knitr::include_graphics("Figures/Figure4.png")
```

@@

```{r, out.width = "70%", fig.cap="**Figure 5. Transmitter phenotypes mapped at single-cell-resolution. ** (A) The cell-type connectome graph with nodes coloured based on neurotransmitter phenotype. (B, C) Morphological rendering of neurons with immunogold labelling for pigment dispersing factor (PDF) neuropeptide, ventral (B) and anterior (C) views. (D) Neurons with immunogold labelling for leucokinin. (E) Neurons with immunogold labelling for allatotropin/orexin. (F) Neurons with immunogold labelling for FVamide. (G) Neurons with immunohistochemically mapped expression of sNPF/RYamide, RGWamide and pedal peptide 2/MLD neuropeptides. (H) Neurons with immunohistochemically mapped expression of achatin and immunogold- or immunohistochemically-mapped (SN-IRP2-FMRF) FMRFamide neuropeptide. (I) Neurons with immunogold (SNMIP-vc) or immunohistochemically mapped expression of myoinhibitory peptide (MIP) expression. (J) Neurons with immunohistochemically mapped expression of serotonin and genetically mapped expression of tryptophan hydroxylase (TrpH), a serotonergic marker. (K) Neurons with genetically mapped expression of vesicular glutamate transporter (VGluT), a glutamatergic marker. (L, M) Neurons with genetically mapped expression of vesicular acetylcholine transporter (VAChT) and choline acetyltransferase (ChAT), cholinergic markers, ventral (L) and anterior (M) views. <br> Figure 5—source data 1. Interactive html file with the network shown in panel A. For an interactive version of the graph see also https://jekelylab.github.io/Platynereis_connectome/."}
#| echo: false
knitr::include_graphics("Figures/Figure5.png")
```

@@

```{r, out.width = "100%", fig.cap="**Figure 4—figure supplement 6. Grouped connectivity graph. ** Cell-type level connectome graph with node horizontal positions proportional to the relative ratio of incoming (postsynaptic) and outgoing (presynpatic) sites. Cell-type names are shown under the nodes. <br> Figure 4—figure supplement 6—source data 1. The network in visNetwork format saved as an R RDS source file, saved with bz2 compression. Can be loaded with read_rds(). For an interactive version of the graph see also https://jekelylab.github.io/Platynereis_connectome/."}
#| echo: false
knitr::include_graphics("Figures/Figure4_fig_suppl6.png")
```

@@
